---
layout: post
title:  "Compute mean for activities - all format"
categories: [jekyll]
tags: [mtus]
---


<span class='newthought'>This post</span> will demonstrate how to simply calcualte mean time of different activities for different format. 

```{r, warning=FALSE, message=FALSE}

# Load the libraries and data
library(plyr)
library(dplyr)
library(tidyr)
library(mtusRlocal)
library(knitr)
library(reshape2)
library(xtable)

dta = read.csv('/Users/giacomovagni/site/motsetchoses/_data/dtaSpells.csv')
```

### 1. Mean of Short PP file 

The `sum` of all activites duration for each individuals for a day must be **1440**. 

```{r}
dta %>% group_by(id, day) %>% summarise(sum(duration))
```

The individual mean can be computed such as : 
```{r}
dta %>% group_by(id, activities) %>% summarise(mean(duration))
```

The aggregate mean by day : 
```{r}
dta %>% group_by(day, activities) %>% summarise(mean(duration))
```

The aggregate mean by activities, regardless of the day : 
```{r}
dta %>% group_by(activities) %>% summarise(mean(duration))
```

### 2. Mean of Long PP file 

For Long PP file 

```{r}
# transform 
dtaPP= dta[rep(1:nrow(dta), dta$duration), ] %>% 
  select(-duration) %>% 
  group_by(id, day) %>% 
  mutate(Time = 1:n()) 

dtaPP %>% group_by(id, day, activities) %>% summarise(time = n()) %>% # transform back to short PP 
  group_by(activities, day) %>% 
  summarise(mean(time))
```

### 3. Mean of Sequence File 


#### 3.1. Aggregate  

The mean time of activities for long PP can be computed by the following codes : 

```{r}
# transform PP to Wide 
dtaSeq = dta[rep(1:nrow(dta), dta$duration), ] %>% 
  select(-duration) %>% 
  group_by(id, day) %>% 
  mutate(Time = 1:n()) %>% # PP 
  arrange(id, day, Time) %>% select(-epnum) %>% spread(Time, activities) # Wide 

# colnames renames  
colnames(dtaSeq) = c('idind', 'day', paste("main", 1:1440, sep = ""))
```

```{r}
N = count(dtaSeq) 
N 
# by act only
dtaSeq %>% select(idind, day, matches('main')) %>% 
  melt(id.vars = c("idind", "day")) %>% group_by(value) %>% 
  summarise(n = n()) %>% # back to PP 
  mutate(time =  (n* 1) / 4 ) %>% # divide by count 
  mutate(TimeClock(time)) 

# by act and days 
dtaSeq %>% select(idind, day, matches('main')) %>% 
  melt(id.vars = c("idind", "day")) %>% group_by(day, value) %>% 
  summarise(n = n()) %>% 
  mutate(time =  (n* 1) / 2) %>% # divide by count 
  mutate(TimeClock(time)) 

# if you haev the full data then group by what you want  
dtaSeq %>% select(idind, day, matches('main')) %>% 
  melt(id.vars = c("idind", "day")) %>% group_by(idind, day, value) %>% 
  summarise(n = n()) %>% 
  group_by(value) %>% summarise( mean(n) ) 
  mutate(TimeClock(time)) 
  
# or simply activities 
  dtaSeq %>% select(idind, day, matches('main')) %>% 
    melt(id.vars = c("idind", "day")) %>% 
    group_by(idind, day, value) %>% 
    summarise(n = n()) %>% 
    group_by(value, day) %>% 
    summarise( mean(n) ) 
  mutate(TimeClock(time)) 
``` 

Alternative when you are unsure of the `count` results 

```{r}
dtaSeq %>% select(idind, day, matches('main')) %>% 
  melt(id.vars = c("idind", "day")) %>% group_by(value) %>% 
    summarise(n = n()) %>% # back to PP 
    mutate(p = n / sum(n)) %>% 
    mutate(time =  1440 * p) %>% # divide by count 
  mutate(TimeClock(time)) 
```



#### 3.2. Individual

From a PP file  

```{r}
head(dtaPP)
```


```{r}
dtaPP_sum = dtaPP %>% 
  group_by(id, day, activities) %>% 
  summarise(n = n() * 1) %>% # multiply by the interval - here is 1 minutes 
  mutate(activities = paste('act_', activities, sep = '')) %>% 
  spread(activities, n, fill = 0) 

dtaPP_sum

dtaPP_sum %>% select(-id, -day) %>% summarise_each(funs(mean))
```

#### 3.3. Weighted Sum and Weighted Mean 


```{r}
# weights of weekdays 
weekendproba = (2 / 7)

# weights - 1 is weekdays 
dtaSeq$day_weights = ifelse(dtaSeq$day %in% "Monday", 1, 1 - weekendproba)

# Weighted mean 
dtaSeq %>% select(idind, day, day_weights, matches('main')) %>% 
  melt(id.vars = c("idind", "day", "day_weights")) %>% 
  group_by(idind, day, day_weights, value) %>% 
  summarise(n = n()) %>% 
  mutate(value = paste('act_', value, sep = '')) %>% 
  spread(value, n, fill = 0) %>% 
  summarise(meanTV = mean(act_TV), weightedmean = weighted.mean(act_TV, day_weights)) 

```

The weighted sum 
```{r}
dtaSeq %>% count(day) %>% mutate(p = n / sum(n))
dtaSeq %>% count(day, wt = day_weights) %>% mutate(p = n / sum(n))
```





